---
title: "New york ride"
author: "yybei"
date: "2022/2/18"
output: 
  html_document:
    css: bootstrap.min.css
    toc: true
    toc_float: true
    code_folding: hide
---

# 1. Introduction
In this project,I'd like to focus on Spatial problem of space & time taxi ride demand and time consumption for each taxi trip in New York.

## 1.1 Set up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 999)
```

```{r setup_13, cache=TRUE, message=FALSE}
library(tidyverse)
library(sf)
library(lubridate)
library(tigris)
library(tidycensus)
library(viridis)
library(riem)
library(gridExtra)
library(knitr)
library(kableExtra)
library(caret)
library(RSocrata)
library(data.table)
library(pacman)
library(geosphere)
library(lubridate)
library(readr)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(corrplot)
library(leaflet)
library(stringr)
library(rgdal)
library(mapview)
library(dplyr)
library(geodist)
library(lubridate)
library(tigris)
library(viridis)
library(riem)
library(gridExtra)
library(knitr)
library(mapview)
library(kableExtra)
library(RSocrata)
library(ggcorrplot)
library(gifski)
library(RSocrata)
library(spatstat)
library(raster)
library(spdep)
library(FNN)
library(gganimate)
library(corrplot)
library(RColorBrewer)
library(ggmap)
library(gmapsdistance)
library(datasets)
library(caret)
library(tree)


plotTheme <- theme(
  plot.title =element_text(size=12),
  plot.subtitle = element_text(size=8),
  plot.caption = element_text(size = 6),
  axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
  axis.text.y = element_text(size = 10),
  axis.title.y = element_text(size = 10),
  # Set the entire chart region to blank
  panel.background=element_blank(),
  plot.background=element_blank(),
  #panel.border=element_rect(colour="#F0F0F0"),
  # Format the grid
  panel.grid.major=element_line(colour="#D0D0D0",size=.2),
  axis.ticks=element_blank())

mapTheme <- theme(plot.title =element_text(size=12),
                  plot.subtitle = element_text(size=8),
                  plot.caption = element_text(size = 6),
                  axis.line=element_blank(),
                  axis.text.x=element_blank(),
                  axis.text.y=element_blank(),
                  axis.ticks=element_blank(),
                  axis.title.x=element_blank(),
                  axis.title.y=element_blank(),
                  panel.background=element_blank(),
                  panel.border=element_blank(),
                  panel.grid.major=element_line(colour = 'transparent'),
                  panel.grid.minor=element_blank(),
                  legend.direction = "vertical", 
                  legend.position = "right",
                  plot.margin = ggplot2::margin(1, 1, 1, 1, 'cm'),
                  legend.key.height = unit(1, "cm"), legend.key.width = unit(0.2, "cm"))


palette5 <- c("#eff3ff","#bdd7e7","#6baed6","#3182bd","#08519c")
palette4 <- c("#D2FBD4","#92BCAB","#527D82","#123F5A")
palette2 <- c("#6a51a3","#bcbddc")

options(tigris_class = "sf")
source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")


setwd("D:\\Spring-Penn\\Capstone\\final\\yy")

#setwd("D:\\Spring-Penn\\Capstone\\in progress\\YebeiYao_NY_taxi-trip-duration-analysis")

#setwd("C:\\Spring-Penn\\Capstone\\in progress\\YebeiYao_NY_taxi-trip-duration-analysis")
```


# 2.Load data
## 2.1 Load census data
Considering taxi ride is a kind of commuting method, the commuting pattern of public transit should be potential factors to taxi trip time consumption and taxi demand around city. So I selected several features that should play an importance to the taxi trip for further model building. 

Those features are census data like population in each census tract of travel time to work by different means of transport method. The race composition and house rent in each tract. Those features are likely to play great infection on taxi trip.


```{r get_census, message=FALSE, warning=FALSE, cache=TRUE, results = 'hide'}
# Install Census API Key
census_api_key("e79f3706b6d61249968c6ce88794f6f556e5bf3d", overwrite = TRUE)

# Import Census Info,especially commmuting data.
tracts16 <-  
  get_acs(geography = "tract", variables = c("B08301_001", #Means Of Transportation To Work,total
                                             "B08301_002", #Means Of Transportation To Work,Car, truck, or van,
                                             "B08301_010", #Means Of Transportation To Work, public transit
                                             "B08301_016", #Means Of Transportation To Work, taxicab
                                             
                #travel time to work ----public transit        
                                            
                                             "B08134_067",  #pub Travel time to work, 30-35 min
                                             "B08134_068",  #pub Travel time to work, 35-45 min
                                             "B08134_069",  #pub Travel time to work, 45-60 min
                                             "B08134_070",  #pub Travel time to work, >60 min
                
                 #travel time to work ----ride(car)        
                                            
                                             "B08134_017",  #car Travel time to work, 30-35 min
                                             "B08134_018",  #car Travel time to work, 35-45 min
                                             "B08134_019",  #car Travel time to work, 45-60 min
                                             "B08134_020",  #car Travel time to work, >60 min
                
                #travel time to work ----ride(taxi)        
                                            
                                             "B08134_117",  #taxi Travel time to work, 30-35 min
                                             "B08134_118",  #taxi Travel time to work, 35-45 min
                                             "B08134_119",  #taxi Travel time to work, 45-60 min
                                             "B08134_120",  #taxi Travel time to work, >60 min
                                             
                                            
                                             "B01003_001",   #Total Population
                                             "B19013_001",   #Median Household Income In The Past 12 Months
                                             "B01002_001",   #Median Age
                                             "C15010_001",   #Bachelor Degree or above
                                             
                #Race                        
                                             "B02001_002", #White people
                                             "B02001_003", #Black Or African American Alone
                                             "B02001_004", #American Indian And Alaska Native Alone
                                             "B02001_005", #Asian
                                             
                                             "B08111_017", #Num_Public Transportation_native
                                             "B08111_018",  #Num_Public Transportation_Foreign born
                
                #rent
                                              "B25059_001", #Upper Contract Rent
                                              "B25057_001", #Lower Contract Rent
                                              "B17005_002", #Income below poverty level
                                              "B16010_003", # less than high school, in labor force
                                              "B16010_016" ), # high school, in labor force
                year=2016, state=36, geometry=T,output = "wide")
                

tracts2016 <- tracts16%>%
  
   rename( Total_means_towork = B08301_001E,
           Car_towork = B08301_002E,
           Public_trans_towork = B08301_010E,
           Taxicab_towork = B08301_016E,
           
     
          pub_traveltime_30_35 = B08134_067E,
          pub_traveltime_35_45 = B08134_068E,
          pub_traveltime_45_60 = B08134_069E,
          pub_traveltime_60 = B08134_070E,
          
          car_traveltime_30_35 = B08134_017E,
          car_traveltime_35_45 = B08134_018E,
          car_traveltime_45_60 = B08134_019E,
          car_traveltime_60 = B08134_020E,
          
          taxi_traveltime_30_35 = B08134_117E,
          taxi_traveltime_35_45 = B08134_118E,
          taxi_traveltime_45_60 = B08134_119E,
          taxi_traveltime_60 = B08134_120E,
          
          
          Total_Pop =  B01003_001E,
          Median_income = B19013_001E,
          Median_age = B01002_001E,
          Bachelor_above = C15010_001E,
          
          White_American = B02001_002E,
          Black_American = B02001_003E,
          American_Indian = B02001_004E,
          Asian = B02001_005E,
          
          public_trans_native = B08111_017E,
          public_trans_foreign = B08111_018E,
          
          upper_rent = B25059_001E,
          lower_rent = B25057_001E,
          income_below_poverty = B17005_002E,
          belowhigh_onemploy = B16010_003E,
          abovehigh_onemploy = B16010_016E)%>%
  
  dplyr::select(Total_means_towork,Car_towork,Public_trans_towork,Taxicab_towork,
         
         pub_traveltime_30_35,pub_traveltime_35_45,pub_traveltime_45_60,pub_traveltime_60,
         car_traveltime_30_35,car_traveltime_35_45,car_traveltime_45_60,car_traveltime_60,
         taxi_traveltime_30_35,taxi_traveltime_35_45,taxi_traveltime_45_60,taxi_traveltime_60,
         
         Total_Pop,Median_income,Median_age,Bachelor_above,
         White_American,Black_American,American_Indian,Asian,
         public_trans_native,public_trans_foreign,
         upper_rent,lower_rent,income_below_poverty,belowhigh_onemploy,abovehigh_onemploy,
         GEOID,NAME,geometry)



clean_trancts_2016 <- tracts2016%>%
    st_transform('ESRI:102318')%>%
  mutate(Pct_publictrans_towork = Public_trans_towork / Total_means_towork,
         Pct_ride_towork = (Car_towork+Taxicab_towork) / Total_means_towork,
         pub_traveltime_30_60 = pub_traveltime_30_35 + pub_traveltime_35_45 + pub_traveltime_45_60,
         ride_traveltime_30_60 = car_traveltime_30_35 + car_traveltime_35_45 + car_traveltime_45_60 +
                                taxi_traveltime_30_35 + taxi_traveltime_35_45 + taxi_traveltime_45_60,
         ride_traveltime_60 = car_traveltime_60 + taxi_traveltime_60,
         pct_bachelor = Bachelor_above / Total_Pop,
         pct_white = White_American / Total_Pop,
         pct_black = Black_American / Total_Pop,
         pct_indian = American_Indian / Total_Pop,
         pct_asian = Asian / Total_Pop,
         pct_poverty = income_below_poverty / Total_Pop,
         pct_onemployment = (belowhigh_onemploy + abovehigh_onemploy) / Total_Pop,
         pct_pubtrans_native = public_trans_native / Public_trans_towork,
         pct_pubtrans_foreign = public_trans_foreign / Public_trans_towork
          )%>%
  dplyr::select (-Total_means_towork,-Car_towork,-Public_trans_towork,-Taxicab_towork,
                 -pub_traveltime_30_35,-pub_traveltime_35_45,-pub_traveltime_45_60,
                 -car_traveltime_30_35,-car_traveltime_35_45,-car_traveltime_45_60,-car_traveltime_60,
                 -taxi_traveltime_30_35,-taxi_traveltime_35_45,-taxi_traveltime_45_60,-taxi_traveltime_60,
                 -Bachelor_above,
                 -White_American,-Black_American,-American_Indian,-Asian,
                 -public_trans_native,-public_trans_foreign,
                 -income_below_poverty,-belowhigh_onemploy,-abovehigh_onemploy
                 )

```

## 2.2 Load base geometry

I'm going to take census tract as the unit of analysis.
And the boundary of the analysis takes the border of New York City.I'm going to filter out the taxi trips happen outside the border later on.

```{r boundary,warning = FALSE, message = FALSE}

NYC_boundary <- st_read("Raw data/NYC Borough Boundaries.geojson")%>% st_transform('ESRI:102318')

clean_trancts_2016_only <- st_intersection(clean_trancts_2016,st_union(NYC_boundary))


NY_Tracts <- 
  clean_trancts_2016_only %>%
  as.data.frame() %>%
  distinct(GEOID, .keep_all = TRUE) %>%
  dplyr::select(GEOID, geometry) %>% 
  st_sf

NY_Tracts <- NY_Tracts %>%
          st_transform(crs = 4326)

```


##  2.3 Load and wrangle taxi trip data
### 2.3.1 Processing taxi trip data
#### Transfer data type
In this section, I imported taxi trip data and made several basic processing work.
The dataset collects various set of information related to taxi trip, including pickup and drop off time and location. Also, it qualifies as real-world data, not only because of the way it has been generated and collected, but also the way it shows, with missing data, unimportant columns, baked-in biases, etc. Thus, before I conduct exploratory analysis, I firstly cleaned the dataset, create some explicit features and got the final meaningful records.

First, I transferred the data column property. 
At beginning, all the columns are in character type, which means although the value shows in numeric or time form, it can not be recognized in specific form. So I transferred longitude, latitude from character into numeric, and the pickup and dropoff time into "timestamp" form.
In this way, I can easily map the start and end location in point geometry form, and extract detailed time property like hour in a day, or weekday in a week for further analysis.

```{r ridership,warning = FALSE, message = FALSE}
taxitrip <- st_read("Raw data/NY-Ridership.csv")


taxitrip$passenger_count <-as.numeric(taxitrip$passenger_count)
taxitrip$trip_duration <-as.numeric(taxitrip$trip_duration)

  taxitrip$vendor_id <-as.numeric(taxitrip$vendor_id)
  taxitrip$dropoff_longitude <-as.numeric(taxitrip$dropoff_longitude)
  taxitrip$dropoff_latitude <-as.numeric(taxitrip$dropoff_latitude)
  taxitrip$pickup_longitude <-as.numeric(taxitrip$pickup_longitude)
  taxitrip$pickup_latitude <-as.numeric(taxitrip$pickup_latitude)
  


taxitrip <- taxitrip %>%
  mutate(pickup_datetime = ymd_hms(pickup_datetime),
         dropoff_datetime = ymd_hms(dropoff_datetime))

```

#### Calculate distance
Calculate distance of each trip. 
```{r distance, message = FALSE, warning = FALSE}
startpoint <- st_as_sf(x = taxitrip, coords = c('pickup_longitude', 'pickup_latitude'), crs = 4326)%>%
  mutate( id = row_number() )
endpoint <- st_as_sf(x = taxitrip, coords = c('dropoff_longitude', 'dropoff_latitude'), crs = 4326)%>%
  mutate( id = row_number() )

start <- taxitrip%>% dplyr::select('pickup_longitude', 'pickup_latitude')
end <-  taxitrip%>% dplyr::select('dropoff_longitude', 'dropoff_latitude')

dis <- geodist_vec(
  start$pickup_longitude,
  start$pickup_latitude,
  end$dropoff_longitude,
  end$dropoff_latitude,
  paired = TRUE,
  sequential = FALSE,
  pad = FALSE,
  measure = "vincenty"
)%>%
  as.data.frame()%>%
  mutate( RowID = row_number() )

taxitrip <- taxitrip%>%mutate( RowID = row_number() )
  
taxitrip <- left_join(taxitrip,dis,by="RowID")%>%
  rename(distance = ".")

```

#### parse time properties
Mutate new columns into taxitrip dataset, to extract time property such as weeks, hours, and weekday.
```{r timeparse, message = FALSE, warning = FALSE}
 taxitrip<- taxitrip %>% 
  mutate(interval60 = floor_date(ymd_hms(pickup_datetime), unit = "hour"),
         week = week(interval60),
         hour = hour(interval60),
         WeekDay = lubridate::wday(interval60, label=TRUE))

  taxitrip$weekend <- ifelse(taxitrip$WeekDay =="周六" |taxitrip$WeekDay == "周日","Weekend","not-Weekend")
```

#### check the current duration
Although the original data set has provided trip duration in each row, however, since the time consumption is the key dependent variable for our prediction model, it is necessary to check if the duration provided is true or false. The safest way is to calculate the time interval between pickup time and dropoff time manually and to make comparison to the existing trip duration, in this way, we can see if the existing trip duration match the calculation result or not. Learning from the checking result, it shows all the duration equals the calculation between pick up and drop off interval.
```{r check, message = FALSE, warning = FALSE}
  taxitrip <- taxitrip %>%
  mutate(duration = as.numeric(abs(int_length(interval(dropoff_datetime,pickup_datetime)))))
  taxitrip$check <- ifelse(taxitrip$trip_duration == taxitrip$duration,TRUE,FALSE)
```

### 2.3.2 Cleaning taxitrip dataset
In this section, I'm going to clean the full taxi trip data set,filter out strange outliers that might affect the accuracy of the prediction. The below statistical summary of the data can help me identify any obvious outliers
```{r cleaning, message = FALSE, warning = FALSE}
kable(summary(taxitrip),format = "html")%>%
  kable_styling(bootstrap_options = c("striped", "hover"))%>%
  kable_classic(full_width = T, html_font = "Cambria")
```

Learning from the above summary table, we can see that some of the feature has shown strange pattern and should be removed for good.
  * **id**: original id is a set of distinct character that represent each individual trips.
  * **vendor_id**: There are 2 unique,from which the minimum value is 1 and the maximum value is 2, which makes sense as we saw there are only two vendor ids 1 and 2.
  * **passenger_count**: There are 9 unique passenger counts.However, the minimum value of passenger count is 0 while the maximum value of passenger count is 10, which is impossible! Firstly,all the trip should have at least one passenger, or the trip will not count, so I should **remove any trip that has none passenger**, where should be some error happens. Secondly, according to the NYC Taxi&Limousine Commission, the maximum number of people, by law is five passengers in one taxicab. Considering the condition that there are passengers under the age of 7 and doesn't take any space and prefer sitting on adults' leg. The maximum passenger of each trip should be 7. So I have to remove any trip record that has passengers **higher than 7**.
  * **Lon and Lat**: No matter the long or lat of pick-up and drop-off location, the trip should be happened inside the boundary of New York City,According to coordinate system, **the Latitude should lie between 40.4775 and 40.9147, while Longitude should lie between - 74.2782 and - 73.7062.** So I have to remove any trips that has those four features locates outside the coordinate range.
  * **store_and_fwd_flag**: This flag indicates whether the trip record was held in vehicle memory before sending to the vendor because the vehicle did not have a connection to the server - Y=store and forward; N=not a store and forward trip.
  * **trip_duration**: duration of the trip in seconds.According to the summary,the maximum duration of trip is 3526282 seconds,converting to hours is 980 hours. If this time is really the trip time consumption, it could be ridiculous. Maybe it's because the owner of the taxi forgot to turn off the meter before he leave the car for a long time. In this case, it will be helpful to **remove any trip that has time duration more than two standard deviations away from the mean duration time**, which means about 95% of values will be within 2 standard deviations of the mean.
  * **distance**: geographic distance between two co-ordinates
  
  After all the cleaning work, by comparison, it shows that I removed overall 3500 observations from the original taxitrip records, which takes over 0.24% of the dataset.
```{r}
#filter out outliers in passenger count.
taxitrip_clean <- filter(taxitrip,passenger_count != 0)
taxitrip_clean <- filter(taxitrip_clean,passenger_count < 8)

#filter out outliers in long and lat.
taxitrip_clean <- filter(taxitrip_clean,pickup_latitude <= 40.9147 & pickup_latitude >= 40.4775)
taxitrip_clean <- filter(taxitrip_clean,dropoff_latitude <= 40.9147 & dropoff_latitude >= 40.4775)
taxitrip_clean <- filter(taxitrip_clean,pickup_longitude <= -73.7062 & pickup_longitude >= -74.2782)
taxitrip_clean <- filter(taxitrip_clean,dropoff_longitude <= -73.7062 & dropoff_longitude >= -74.2782)

#filter out outliers in trip duration.
mean(taxitrip_clean$trip_duration)
sd(taxitrip_clean$trip_duration)
taxitrip_clean <- filter(taxitrip_clean,trip_duration <= mean(trip_duration)+2*sd(trip_duration) & trip_duration >= mean(trip_duration)-2*sd(trip_duration))

record_Removed <- nrow(taxitrip)-nrow(taxitrip_clean)
record_Removed_pct <- (nrow(taxitrip)-nrow(taxitrip_clean))/nrow(taxitrip)
paste(record_Removed_pct*100, "%", sep='')


taxitrip_clean <- taxitrip_clean %>%
        mutate(time_of_day = case_when(hour(interval60) < 7 | hour(interval60) > 18 ~ "Overnight",
                                 hour(interval60) >= 7 & hour(interval60) < 10 ~ "Morning Rush Hour",
                                 hour(interval60) >= 10 & hour(interval60) < 15 ~ "Mid-Day",
                                 hour(interval60) >= 15 & hour(interval60) <= 18 ~ "Afternoon Rush Hour"))
```

Summary of cleaned trip data.
```{r cleaning2, message = FALSE, warning = FALSE}
kable(summary(taxitrip_clean),format = "html")%>%
  kable_styling(bootstrap_options = c("striped", "hover"))%>%
  kable_classic(full_width = T, html_font = "Cambria")
```
### 2.3.3 Exploratory data analysis of taxitrip
#### Create sample taxi trip subset
Since the overall taxi trip record is too big, I create a random sample for exploratory data analysis that hard to visualize for large dataset.
```{r}
# Random sample subset of trip data for exploratory data analysis, set seed to be reproduced and create sample trip data.
set.seed(123)
taxitrip_sample <- sample_n(taxitrip_clean, 100000)

```

#### Trip data sample mapping
Observing from the map,it shows point geometry scattered around city, it is obvious that Manhattan Island and Brooklyn of NYC burdens pretty much higher taxi ride demand. And drop off locations spreads farther than pick up locations.
```{r}
dat_amenity_dropoff <- dat_census_cleanoutliers_amenity%>%
  st_as_sf(coords = c("dropoff_longitude","dropoff_latitude"), crs = 4326, agr = "constant")%>%
  drop_na()

dat_amenity_pickup <- dat_census_cleanoutliers_amenity%>%
  st_as_sf(coords = c("pickup_longitude","pickup_latitude"), crs = 4326, agr = "constant")%>%drop_na()



dropoff <- ggplot()+
  geom_sf(data = NY_Tracts)+
  geom_sf(data = dat_amenity_dropoff, show.legend = "point", size= .5, alpha = 0.5,color="#df65b0") +
  labs(title = "Taxi trip general map - dropoff location") +
  mapTheme() + 
  theme(plot.title = element_text(size=20))

pickup <- ggplot()+
  geom_sf(data = NY_Tracts)+
  geom_sf(data = dat_amenity_pickup, show.legend = "point", size= .5, alpha = 0.5,color="#3182bd") +
  labs(title = "Taxi trip general map - pickup location") +
  mapTheme() + 
  theme(plot.title = element_text(size=20))

grid.arrange(dropoff,pickup, nrow = 1)

```


kable(summary(weather.Data.clean),format = "html")%>%
  kable_styling(bootstrap_options = c("striped", "hover"))%>%
  kable_classic(full_width = T, html_font = "Cambria")






#### Univariate Analysis 
  + For target variable **trip_duration**
Learning from the histogram, the count of trip duration shows a long skew as the time consumption goes up. But if I use log transfermation to the trip duration, it shows a normal distribution pattern.
```{r}
original_his <-dat_census_cleanoutliers_amenity %>%
  ggplot(aes(trip_duration)) +
  geom_histogram(color = "#756bb1",fill = "#f1eef6", bins = 150) +
    plotTheme()


log_his <- dat_census_cleanoutliers_amenity %>%
  ggplot(aes(trip_duration)) +
  geom_histogram(color = "#756bb1",fill = "#f1eef6", bins = 150) +
  scale_x_log10() +
  scale_y_sqrt()+
    plotTheme()

grid.arrange(original_his,log_his, nrow = 1)
```

  + For target variable **passenger_count**
Since the box plot of trip duration originally shows toooo many outliers, I made comparison between 
```{r}
normal_passct <- ggplot(dat_census_cleanoutliers_amenity,aes(x=as.factor(passenger_count), y=trip_duration, color = passenger_count))+
geom_boxplot() +
    theme(legend.position = "none") +
    labs(title = "Trip Duration by Number of Passengers", x = "Number of passengers",y= "Trip duration")+
    plotTheme()


log_passct <- ggplot(dat_census_cleanoutliers_amenity,aes(x=as.factor(passenger_count), y=trip_duration, color = passenger_count))+
geom_boxplot() +
  scale_y_log10() + 
  theme(legend.position = "none") +
    labs(title = "Log of Trip Duration by Number of Passengers", x = "Number of passengers",y= "Trip duration")+
    plotTheme()


grid.arrange(normal_passct,log_passct, nrow = 2)
```


  + For target variable **store_and_fwd_flag**
  This is a binary variable indicates whether the trip record was held in vehicle memory before sending to the vendor because the vehicle did not have a connection to the server,where Y=store and forward; N=not a store and forward trip.
```{r}
taxitrip_sample %>% 
  group_by(store_and_fwd_flag) %>% 
  count() %>% 
  ggplot(aes(x=store_and_fwd_flag,y=n, fill=store_and_fwd_flag))+ 
  geom_col()+
  scale_fill_manual(values = palette2)+
    plotTheme()
```

 + For target variable **Check or not**
  This is a binary variable indicates whether the trip duration provided by the original data set is true or not.Learning from the plot, it is **one hundred percent** that the trip duration is correct with the calculation from the time stamp.
```{r}
taxitrip_sample %>% 
  group_by(check) %>% 
  count() %>% 
  ggplot(aes(x=check,y=n, fill=check))+ 
  geom_col()+
  scale_fill_manual(values = palette2)+
    plotTheme()
```
  + For target variable **distance and weekday in a week**
  Here I plot both the sum of distance across the week day, and the mean value of distance of a week.I thought the pattern of two plots should look similar, but result is surprising! As plots shown below, especially in the cases of Monday, Thursday and Friday. On Monday, the sum of distance is the lowerest, but the mean value of the distance is second highest, which means the trip count is least but people tend to take taxi in longer ridership, maybe take taxi from outside town to center city, etc.
  As on Thursday and Friday, the mean value of distance is in the middle, while the sum of distance are the highest. Which means during these days, people tend to take taxi to relatively near place, with lower price but more frequent call. Maybe it's becuase people tend to hang out on these days.In other words, the distribution pattern of distance and trip count looks similar all the same.
```{r fig.height=6, fig.width=15}

dis <- taxitrip_sample%>%
   dplyr::select(WeekDay,distance)%>%
  group_by(WeekDay)  %>%
   summarise(dis = sum(distance))

meandis<-taxitrip_sample%>%
   dplyr::select(WeekDay,distance)%>%
  group_by(WeekDay)  %>%
   summarise(meandis = mean(distance))

displot <- ggplot(dis,aes(WeekDay,y=dis)) +
  geom_bar(stat="identity", aes(fill = dis))+
  scale_fill_gradient(low="#bdd7e7", high='darkblue')+
  labs(x = 'Day of a week', y = 'Sum of distance', title = 'Distance by day of a week')+
    plotTheme()

meandisplot <- ggplot(meandis,aes(WeekDay,y=meandis)) +
  geom_bar(stat="identity", aes(fill = meandis))+
  scale_fill_gradient(low="#bdd7e7", high='darkblue')+
  labs(x = 'Day of a week', y = 'Mean distance', title = 'Mean distance by day of a week')+
    plotTheme()

count <- taxitrip_sample%>%
   dplyr::select(WeekDay)%>%
  group_by(WeekDay)  %>%
   count()

countplot <- ggplot(count,aes(WeekDay,y=n)) +
  geom_bar(stat="identity", aes(fill = n))+
  scale_fill_gradient(low="#bdd7e7", high='darkblue')+
  labs(x = 'Day of a week', y = 'trip count', title = 'Mean distance by day of a week')+
    plotTheme()


grid.arrange(displot,meandisplot, ncol = 2)
```

  + For target variable **Hour in a day**
  Here I plot both the mean trip duration and trip count across each 24 hours in a day.We can see the peak of trip count is at noon to afternoon, while the peak of trip distance is in late afternoon. Although in midnight,the trip count is low, the average trip duration is similar to other times.
```{r fig.height=8, fig.width=10}
meandur <- taxitrip_sample%>%
   dplyr::select(hour,trip_duration)%>%
  group_by(hour)%>%
  summarize(meandur=mean(trip_duration))

count <- taxitrip_sample%>%
   dplyr::select(hour)%>%
  group_by(hour)%>%
  count()



 countplot_wk<- ggplot(count,aes(x = hour, y = n)) + 
   geom_bar(stat = 'identity',aes(fill=n)) +
 coord_polar(start = 0)  +
   scale_fill_gradient(low="#bdd7e7", high='darkblue')+
 ylab("Count") + ggtitle("Trip Count by hour in a day") + 
    scale_x_continuous("", limits = c(0, 24), breaks = seq(0, 24), labels = seq(0,24))


 meandurplot_wk <- ggplot(meandur,aes(x = hour, y = meandur)) + 
   geom_bar(stat = 'identity',aes(fill=meandur)) +
 coord_polar(start = 0)  +
   scale_fill_gradient(low="#bdd7e7", high='darkblue')+
 ylab("Mean duration") + ggtitle("Mean of trip duration by hour in a day") + 
    scale_x_continuous("", limits = c(0, 24), breaks = seq(0, 24), labels = seq(0,24))

grid.arrange(countplot_wk,meandurplot_wk, ncol = 2)
```
  + For target variable **time_of_day**
By classifying one day into four main periods,the distance and trip count distribution across day is clear.It is surprising that during both morning and afternoon rush hour,trip distance is lower than other time. Also, the trip count during morning rush hour is the lowest one! Which means during rush hour, people choose to commute to work in other ways instead of taking taxi.
  
```{r fig.height=6, fig.width=15}
timeofday_disgroup <- taxitrip_sample%>%
  dplyr::select(time_of_day,distance)%>%
  group_by(time_of_day)%>%
  summarize(dis_time=mean(distance))

timeofday_disgroupplot <- 
  ggplot(timeofday_disgroup,aes(x =time_of_day, y = dis_time))+
   geom_bar(stat = 'identity',aes(fill=dis_time)) +
   scale_fill_gradient(low="#bdd7e7", high='darkblue')+
  labs(x = 'Time of Day', y = 'Mean Travel Distance', title = 'Trip Distance,Four Period of Time in a Day')+
    plotTheme()


timeofday_countgroup <- taxitrip_sample%>%
  dplyr::select(time_of_day)%>%
  group_by(time_of_day)%>%
  count()

timeofday_countgroupplot <-
ggplot(timeofday_countgroup,aes(x =time_of_day, y = n))+
   geom_bar(stat = 'identity',aes(fill=n)) +
   scale_fill_gradient(low="#bdd7e7", high='darkblue')+
  labs(x = 'Time of Day', y = 'Trip Count', title = 'Trip Count,Four Period of Time in a Day')+
    plotTheme()

grid.arrange(timeofday_disgroupplot,timeofday_countgroupplot, ncol = 2)
```

# 3.External Data
## 3.1 Import meteorological data
With the purpose of accurately predict the taxi demand spatially and timely by selected affecting factors, the impact of air quality, wind speed, temperature and wind speed should be taken into consideration.With "riem_measures" package through IBM bought Weather Company and Weather Underground API, I extract temperature, precipitation, and wind speed for every hour between January and June in 2016, in NYC station
```{r fig.height=6,fig.width=6}
weather.Data <- 
  riem_measures(station = "NYC", date_start = "2016-01-01", date_end = "2016-06-30")

weather.Data.clean <- weather.Data%>%
  dplyr::select(valid, tmpf, p01i, sknt)%>%
  replace(is.na(.), 0) %>%
    mutate(interval60 = ymd_h(substr(valid,1,13))) %>%
    mutate(week = week(interval60),
           dotw = lubridate::wday(interval60, label=TRUE)) %>%
    group_by(interval60) %>%
    summarize(Temperature = max(tmpf),
              Precipitation = sum(p01i),
              Wind_Speed = max(sknt)) %>%
    mutate(Temperature = ifelse(Temperature == 0, 42, Temperature))

grid.arrange(top = "Weather Data - NYC - Jan & June, 2016",
  ggplot(weather.Data.clean, aes(interval60,Precipitation)) + geom_line(colour = "#045a8d") + 
    labs(title="Percipitation", x="Hour", y="Percipitation") + plotTheme(),
  ggplot(weather.Data.clean, aes(interval60,Wind_Speed)) + geom_line(colour = "#045a8d") + 
    labs(title="Wind Speed", x="Hour", y="Wind Speed") + plotTheme(),
  ggplot(weather.Data.clean, aes(interval60,Temperature)) + geom_line(colour = "#045a8d") + 
    labs(title="Temperature", x="month", y="Temperature") + plotTheme())

```


## 3.2 Import Other data
## 3.3 Come to the final panel
```{r}
sf_use_s2(FALSE)


trip_join_census <- st_join(taxitrip_sample %>% 
          filter(is.na(pickup_latitude) == FALSE &
                   is.na(pickup_longitude) == FALSE &
                   is.na(dropoff_latitude) == FALSE &
                   is.na(dropoff_longitude) == FALSE) %>%
          st_as_sf(., coords = c( "pickup_longitude","pickup_latitude"), crs = 4326),
        NY_Tracts,
        join=st_intersects,
              left = TRUE) %>%
  rename(pickup.Tract = GEOID) %>%
  mutate(pickup_longitude = unlist(map(geometry, 1)),
         pickup_latitude = unlist(map(geometry, 2)))%>%
  as.data.frame() %>%
  dplyr::select(-geometry)%>%
  st_as_sf(., coords = c( "dropoff_longitude", "dropoff_latitude"), crs =4326) %>%
  st_join(., NY_Tracts,
          join=st_intersects,
          left = TRUE) %>%
  rename(dropoff.Tract = GEOID)  %>%
  mutate(dropoff_longitude = unlist(map(geometry, 1)),
         dropoff_latitude = unlist(map(geometry, 2)))%>%
  as.data.frame() %>%
  dplyr::select(-geometry)

trip_join_census<-trip_join_census%>%
  dplyr::select(-duration,-check)

```

```{r panel_length_check , message = FALSE, warning = FALSE}
length(unique(trip_join_census$interval60)) * length(unique(trip_join_census$pickup.Tract))


study.panel <- 
  expand.grid(interval60=unique(trip_join_census$interval60), 
              pickup.Tract = unique(trip_join_census$pickup.Tract)) %>%
  left_join(., trip_join_census %>%
              dplyr::select(pickup.Tract,pickup_longitude, pickup_latitude )%>%
              distinct() %>%
              group_by(pickup.Tract) %>%
              slice(1))



nrow(study.panel)      
```

```{r}
ride.panel2 <- 
  trip_join_census %>%
  mutate(Trip_Counter = 1) %>%
  right_join(study.panel) %>% 
  group_by(interval60, pickup.Tract, pickup_longitude, pickup_latitude) %>%
  summarize(Trip_Count = sum(Trip_Counter, na.rm=T)) %>%
  left_join(weather.Data.clean) %>%
  ungroup() %>%
  filter(is.na(pickup.Tract) == FALSE)
```

```{r}
ride.panel2 <- 
  left_join(ride.panel2,NY_Tracts %>%
              as.data.frame() %>%
              dplyr::select(-geometry), by = c("pickup.Tract" = "GEOID"))
```


```{r}
ride.panel2 <- 
  ride.panel2 %>% 
  arrange(pickup.Tract, interval60) %>% 
  mutate(lagHour = dplyr::lag(Trip_Count,1),
         lag2Hours = dplyr::lag(Trip_Count,2),
         lag3Hours = dplyr::lag(Trip_Count,3),
         lag4Hours = dplyr::lag(Trip_Count,4),
         lag12Hours = dplyr::lag(Trip_Count,12),
         lag1day = dplyr::lag(Trip_Count,24))
```



```{r}
as.data.frame(ride.panel2) %>%
    group_by(interval60) %>% 
    summarise_at(vars(starts_with("lag"), "Trip_Count"), mean, na.rm = TRUE) %>%
    gather(Variable, Value, -interval60, -Trip_Count) %>%
    mutate(Variable = factor(Variable, levels=c("lagHour","lag2Hours","lag3Hours","lag4Hours",
                                                "lag12Hours","lag1day")))%>%
    group_by(Variable) %>%  
    summarize(correlation = round(cor(Value, Trip_Count),2))%>%
   kable(caption = "Time Lag Correlation") %>%
    kable_styling("striped", full_width = F)
```
```{r}
panel <- 
  trip_join_census %>%
  left_join(ride.panel2) %>%
  filter(is.na(pickup.Tract) == FALSE) %>%
  filter(is.na(dropoff.Tract) == FALSE)

panel<- 
  inner_join(panel, clean_trancts_2016_only %>%
              as.data.frame()%>%
              dplyr::select(-geometry), by = c("pickup.Tract" = "GEOID"))


count.panel <-
  trip_join_census %>%
  left_join(ride.panel2) %>%
  filter(is.na(pickup.Tract) == FALSE) %>%
  filter(is.na(dropoff.Tract) == FALSE)


```


# Correlation between trip duration and features
```{r}
panel%>%
  dplyr::select(-RowID,-id,-pickup_latitude,-pickup_longitude,-pickup_datetime,-dropoff_latitude,-dropoff_longitude,-dropoff_datetime,-pickup.Tract,-dropoff.Tract,-vendor_id)%>%
  gather(Variable, Value, -trip_duration) %>% 
   ggplot(aes(Value, trip_duration)) +
     geom_point(size = .5) + geom_smooth(method = "lm", se=F, colour = "#FA7800") +
     facet_wrap(~Variable, ncol = 3, scales = "free") +
     labs(title = "trip duration as a function of continuous variables") +
     plotTheme()


numericVars <- 
  select_if(panel, is.numeric) %>% na.omit()%>%
  dplyr::select(-RowID,-pickup_latitude,-pickup_longitude,-dropoff_latitude,-dropoff_longitude)


ggcorrplot(
  round(cor(numericVars), 1), 
  p.mat = cor_pmat(numericVars),
  colors = c("#88419d", "#fff7ec", "#08519c"),
  type="lower",
  insig = "blank") +  
    labs(title = "Correlation across numeric variables") 

```
















```{r}
 

# Fitting a tree to lat and long
fitted_tree <- tree(formula = trip_duration  ~ passenger_count + store_and_fwd_flag + distance + hour + WeekDay + weekend + time_of_day + Median_income+upper_rent+
       Temperature + Precipitation + Wind_Speed,
                    data = trip.training)

# Draw a diagram of the tree structure
plot(fitted_tree)
text(fitted_tree)

summary(fitted_tree)
```


