---
title: "Capstone-datawrangling-yy0208"
author: "yybei"
date: "2022/2/8"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r setup_13, cache=TRUE, message=FALSE}
library(tidyverse)
library(sf)
library(lubridate)
library(tigris)
library(tidycensus)
library(viridis)
library(riem)
library(gridExtra)
library(knitr)
library(kableExtra)
library(RSocrata)
library(osmdata)
library(sp)
library(rgdal)
library(rgeos)
library(spdep)
library(spgwr)
library(tmap)

plotTheme <- theme(
  plot.title =element_text(size=12),
  plot.subtitle = element_text(size=8),
  plot.caption = element_text(size = 6),
  axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
  axis.text.y = element_text(size = 10),
  axis.title.y = element_text(size = 10),
  # Set the entire chart region to blank
  panel.background=element_blank(),
  plot.background=element_blank(),
  #panel.border=element_rect(colour="#F0F0F0"),
  # Format the grid
  panel.grid.major=element_line(colour="#D0D0D0",size=.2),
  axis.ticks=element_blank())

mapTheme <- theme(plot.title =element_text(size=12),
                  plot.subtitle = element_text(size=8),
                  plot.caption = element_text(size = 6),
                  axis.line=element_blank(),
                  axis.text.x=element_blank(),
                  axis.text.y=element_blank(),
                  axis.ticks=element_blank(),
                  axis.title.x=element_blank(),
                  axis.title.y=element_blank(),
                  panel.background=element_blank(),
                  panel.border=element_blank(),
                  panel.grid.major=element_line(colour = 'transparent'),
                  panel.grid.minor=element_blank(),
                  legend.direction = "vertical", 
                  legend.position = "right",
                  plot.margin = margin(1, 1, 1, 1, 'cm'),
                  legend.key.height = unit(1, "cm"), legend.key.width = unit(0.2, "cm"))

palette5 <- c("#eff3ff","#bdd7e7","#6baed6","#3182bd","#08519c")
palette4 <- c("#D2FBD4","#92BCAB","#527D82","#123F5A")
palette2 <- c("#6baed6","#08519c")

qBr <- function(df, variable, rnd) {
  if (missing(rnd)) {
    as.character(quantile(round(df[[variable]],0),
                          c(.01,.2,.4,.6,.8), na.rm=T))
  } else if (rnd == FALSE | rnd == F) {
    as.character(formatC(quantile(df[[variable]]), digits = 3),
                 c(.01,.2,.4,.6,.8), na.rm=T)
  }
}

q5 <- function(variable) {as.factor(ntile(variable, 5))}
```

Load census API key

```{r install_census_API_key, warning = FALSE, include=FALSE, eval = TRUE}
# Install Census API Key
census_api_key("e79f3706b6d61249968c6ce88794f6f556e5bf3d", overwrite = TRUE)
```

set dictonary and boundary
```{r}
setwd("D:\\Spring-Penn\\Capstone\\in progress\\0209")

chicagoBoundary <- 
  st_read("ChicagoBoundaries - City.geojson") %>%
  st_transform('ESRI:102271')
```
Import Census Info
```{r get_census, message=FALSE, warning=FALSE, cache=TRUE, results = 'hide'}
tracts19 <-  
  get_acs(geography = "tract", variables = c("B08301_001", #Means Of Transportation To Work,total
                                             "B08301_002", #Means Of Transportation To Work,Car, truck, or van,
                                             "B08301_010", #Means Of Transportation To Work, public transit
                                             "B08301_016", #Means Of Transportation To Work, taxicab
                                             
                #travel time to work ----public transit        
                                            
                                             "B08134_067",  #pub Travel time to work, 30-35 min
                                             "B08134_068",  #pub Travel time to work, 35-45 min
                                             "B08134_069",  #pub Travel time to work, 45-60 min
                                             "B08134_070",  #pub Travel time to work, >60 min
                
                 #travel time to work ----ride(car)        
                                            
                                             "B08134_017",  #car Travel time to work, 30-35 min
                                             "B08134_018",  #car Travel time to work, 35-45 min
                                             "B08134_019",  #car Travel time to work, 45-60 min
                                             "B08134_020",  #car Travel time to work, >60 min
                
                #travel time to work ----ride(taxi)        
                                            
                                             "B08134_117",  #taxi Travel time to work, 30-35 min
                                             "B08134_118",  #taxi Travel time to work, 35-45 min
                                             "B08134_119",  #taxi Travel time to work, 45-60 min
                                             "B08134_120",  #taxi Travel time to work, >60 min
                                             
                                            
                                             "B01003_001",   #Total Population
                                             "B19013_001",   #Median Household Income In The Past 12 Months
                                             "B01002_001",   #Median Age
                                             "C15010_001",   #Bachelor Degree or above
                                             
                #Race                        
                                             "B02001_002", #White people
                                             "B02001_003", #Black Or African American Alone
                                             "B02001_004", #American Indian And Alaska Native Alone
                                             "B02001_005", #Asian
                                             
                                             "B08111_017", #Num_Public Transportation_native
                                             "B08111_018",  #Num_Public Transportation_Foreign born
                
                #rent
                                              "B25059_001", #Upper Contract Rent
                                              "B25057_001", #Lower Contract Rent
                                              "B17005_002", #Income below poverty level
                                              "B16010_003", # less than high school, in labor force
                                              "B16010_016" ), # high school, in labor force
                year=2019, state=17, county=031, geometry=T,output = "wide")
                

tracts2019 <- tracts19%>%
  
   rename( Total_means_towork = B08301_001E,
           Car_towork = B08301_002E,
           Public_trans_towork = B08301_010E,
           Taxicab_towork = B08301_016E,
           
     
          pub_traveltime_30_35 = B08134_067E,
          pub_traveltime_35_45 = B08134_068E,
          pub_traveltime_45_60 = B08134_069E,
          pub_traveltime_60 = B08134_070E,
          
          car_traveltime_30_35 = B08134_017E,
          car_traveltime_35_45 = B08134_018E,
          car_traveltime_45_60 = B08134_019E,
          car_traveltime_60 = B08134_020E,
          
          taxi_traveltime_30_35 = B08134_117E,
          taxi_traveltime_35_45 = B08134_118E,
          taxi_traveltime_45_60 = B08134_119E,
          taxi_traveltime_60 = B08134_120E,
          
          
          Total_Pop =  B01003_001E,
          Median_income = B19013_001E,
          Median_age = B01002_001E,
          Bachelor_above = C15010_001E,
          
          White_American = B02001_002E,
          Black_American = B02001_003E,
          American_Indian = B02001_004E,
          Asian = B02001_005E,
          
          public_trans_native = B08111_017E,
          public_trans_foreign = B08111_018E,
          
          upper_rent = B25059_001E,
          lower_rent = B25057_001E,
          income_below_poverty = B17005_002E,
          belowhigh_onemploy = B16010_003E,
          abovehigh_onemploy = B16010_016E)%>%
  
  select(Total_means_towork,Car_towork,Public_trans_towork,Taxicab_towork,
         
         pub_traveltime_30_35,pub_traveltime_35_45,pub_traveltime_45_60,pub_traveltime_60,
         car_traveltime_30_35,car_traveltime_35_45,car_traveltime_45_60,car_traveltime_60,
         taxi_traveltime_30_35,taxi_traveltime_35_45,taxi_traveltime_45_60,taxi_traveltime_60,
         
         Total_Pop,Median_income,Median_age,Bachelor_above,
         White_American,Black_American,American_Indian,Asian,
         public_trans_native,public_trans_foreign,
         upper_rent,lower_rent,income_below_poverty,belowhigh_onemploy,abovehigh_onemploy,
         GEOID,NAME,geometry)



clean_trancts_2019 <- tracts2019%>%
    st_transform('ESRI:102271')%>%
  mutate(Pct_publictrans_towork = Public_trans_towork / Total_means_towork,
         Pct_ride_towork = (Car_towork+Taxicab_towork) / Total_means_towork,
         pub_traveltime_30_60 = pub_traveltime_30_35 + pub_traveltime_35_45 + pub_traveltime_45_60,
         ride_traveltime_30_60 = car_traveltime_30_35 + car_traveltime_35_45 + car_traveltime_45_60 +
                                taxi_traveltime_30_35 + taxi_traveltime_35_45 + taxi_traveltime_45_60,
         ride_traveltime_60 = car_traveltime_60 + taxi_traveltime_60,
         pct_bachelor = Bachelor_above / Total_Pop,
         pct_white = White_American / Total_Pop,
         pct_black = Black_American / Total_Pop,
         pct_indian = American_Indian / Total_Pop,
         pct_asian = Asian / Total_Pop,
         pct_poverty = income_below_poverty / Total_Pop,
         pct_onemployment = (belowhigh_onemploy + abovehigh_onemploy) / Total_Pop,
         pct_pubtrans_native = public_trans_native / Public_trans_towork,
         pct_pubtrans_foreign = public_trans_foreign / Public_trans_towork
          )%>%
  dplyr::select (-Total_means_towork,-Car_towork,-Public_trans_towork,-Taxicab_towork,
                 -pub_traveltime_30_35,-pub_traveltime_35_45,-pub_traveltime_45_60,
                 -car_traveltime_30_35,-car_traveltime_35_45,-car_traveltime_45_60,-car_traveltime_60,
                 -taxi_traveltime_30_35,-taxi_traveltime_35_45,-taxi_traveltime_45_60,-taxi_traveltime_60,
                 -Bachelor_above,
                 -White_American,-Black_American,-American_Indian,-Asian,
                 -public_trans_native,-public_trans_foreign,
                 -income_below_poverty,-belowhigh_onemploy,-abovehigh_onemploy
                 )


clean_trancts_2019_only <- st_intersection(clean_trancts_2019,chicagoBoundary)
```

ggplot() + 
  geom_sf(data=clean_trancts_2019_only)


```{r extract_geometries }
chicagoTracts <- 
  clean_trancts_2019 %>%
  as.data.frame() %>%
  distinct(GEOID, .keep_all = TRUE) %>%
  na.omit()%>%
  select(GEOID, geometry,NAME) %>% 
  st_sf
```

load taxi ride data in 2019
```{r}
 taxi_ride <- st_read("https://data.cityofchicago.org/resource/h4cq-z3dy.geojson")
```



load facility data
```{r}
school <- st_read("schools.geojson")


hospital <- st_read("Hospitals/Hospitals.shp")

openspace_neigh_garden <- st_read("Openspaces_Neighborspace_Gardens/DATA_ADMIN_OPNSP_NEIGHBORSPACE_GARDENS.shp")

grocery_store <- st_read("https://data.cityofchicago.org/resource/53t8-wyrc.geojson")



hospital.sf <- hospital%>%
  select(geometry)%>%
  na.omit()%>%
  distinct()%>%
  st_transform('ESRI:102271')

school.sf <- school%>%
  select(geometry)%>%
  na.omit()%>%
  distinct()%>%
  st_transform('ESRI:102271')

openspace_neigh_garden.sf <- openspace_neigh_garden%>%
  select(geometry)%>%
  na.omit()%>%
  distinct()%>%
  st_transform('ESRI:102271')

grocery_store.sf <- grocery_store%>%
  select(geometry)%>%
  na.omit()%>%
  distinct()%>%
  st_transform('ESRI:102271')

```


```{r}
# for calculating average nearest neighbor distance.
nn_function <- function(measureFrom,measureTo,k) {
  measureFrom_Matrix <- as.matrix(measureFrom)
  measureTo_Matrix <- as.matrix(measureTo)
  nn <-   
    get.knnx(measureTo, measureFrom, k)$nn.dist
  output <-
    as.data.frame(nn) %>%
    rownames_to_column(var = "thisPoint") %>%
    gather(points, point_distance, V1:ncol(.)) %>%
    arrange(as.numeric(thisPoint)) %>%
    group_by(thisPoint) %>%
    dplyr::summarize(pointDistance = mean(point_distance)) %>%
    arrange(as.numeric(thisPoint)) %>% 
    dplyr::select(-thisPoint) %>%
    pull() # pull() is similar to $. It's mostly useful because it looks a little nicer in pipes, it also works with remote data frames, and it can optionally name the output.
  
  return(output)  
}


```

ggplot() + 
  geom_sf(data=clean_trancts_2019_only)+
  geom_sf(data=hospital.sf)
  
  
```{r}
### Queen weight metrix
queen<-poly2nb(clean_trancts_2019_only, row.names=chicagoTracts$GEOID)
summary(queen)


### Global Moran's I
queenlist<-nb2listw(queen, style = 'W')
moran(clean_trancts_2019_only$pub_traveltime_60, queenlist, n=length(queenlist$neighbours), S0=Szero(queenlist))$`I` #-- Morab's I value for the sample 

moranMC<-moran.mc(clean_trancts_2019_only$pub_traveltime_60, queenlist, nsim=999)  #We use 999 permutations
moranMC #-- p-value

moranMCres_traveltime_morethan_60min<-moranMC$res
hist(moranMCres_traveltime_morethan_60min, freq=10000000, nclass=100)   #-- Draws distribution of Moran's I's calculated from randomly permuted values
abline(v=moran(clean_trancts_2019_only$pub_traveltime_60, queenlist, n=length(queenlist$neighbours), S0=Szero(queenlist))$`I`, col='red')


```
```{r}
moranSig.plot<-function(df,listw, title){
  local<-localmoran(x=df$pub_traveltime_60, listw=listw, zero.policy = FALSE)
  moran.map<-cbind(df, local)
  tm<-tm_shape(moran.map)+
    tm_borders(col='white')+
    tm_fill(col='Pr.z....E.Ii..', style='fixed', breaks=c(0,0.05, 1), title= 'p-value', palette = '-BuPu')+
    tm_layout(frame = FALSE, title = title)
  print(tm)
}
moranSig.plot(clean_trancts_2019_only, queenlist, 'p-value') #-- screenshot the output
```
```{r}
reg<-lm(formula=pub_traveltime_60 ~ Pct_publictrans_towork + pct_bachelor  + pct_white + pct_indian + pct_asian
        +pct_poverty + pct_onemployment + pct_pubtrans_native + pct_pubtrans_foreign + upper_rent + lower_rent, data=clean_trancts_2019_only)
summary(reg) # --copy the result in text
AIC(reg)
logLik(reg)
standardised<-rstandard(reg)
```


Pct_publictrans_towork = Public_trans_towork / Total_means_towork,
         Pct_ride_towork = (Car_towork+Taxicab_towork) / Total_means_towork,
         pub_traveltime_30_60 = pub_traveltime_30_35 + pub_traveltime_35_45 + pub_traveltime_45_60,
         ride_traveltime_30_60 = car_traveltime_30_35 + car_traveltime_35_45 + car_traveltime_45_60 +
                                taxi_traveltime_30_35 + taxi_traveltime_35_45 + taxi_traveltime_45_60,
         ride_traveltime_60 = car_traveltime_60 + taxi_traveltime_60,
         pct_bachelor = Bachelor_above / Total_Pop,
         pct_white = White_American / Total_Pop,
         pct_black = Black_American / Total_Pop,
         pct_indian = American_Indian / Total_Pop,
         pct_asian = Asian / Total_Pop,
         pct_poverty = income_below_poverty / Total_Pop,
         pct_onemployment = (belowhigh_onemploy + abovehigh_onemploy) / Total_Pop,
         pct_pubtrans_native = public_trans_native / Public_trans_towork,
         pct_pubtrans_foreign



```{r}
ggplot(clean_trancts_2019_only)+
  geom_sf(data = chicagoBoundary)+
  geom_sf(aes(fill = q5(pub_traveltime_60)), lwd = 0) +
  scale_fill_manual(values = palette5,
                    labels = qBr(clean_trancts_2019_only, "pubtrans_60min"),
                    name = "pubtransit_under60min\n(Quintile Breaks)") +
  labs(title = "pubtrans_60min") +
  mapTheme() + 
  theme(plot.title = element_text(size=22))


ggplot(clean_trancts_2019_only, aes(fill = pub_traveltime_30_60)) +
    geom_sf()+
    scale_fill_gradientn(colours=rev(magma(5)))
```

```{r}
 
        

bw<-gwr.sel(formula=pub_traveltime_6 ~ Pct_publictrans_towork + pct_bachelor  + pct_white + pct_indian + pct_asian
        +pct_poverty + pct_onemployment + pct_pubtrans_native + pct_pubtrans_foreign + upper_rent + lower_rent,
            data=clean_trancts_2019_only,
            method = "aic",
            adapt = TRUE)
bw

gwrmodel<-gwr(formula=pub_traveltime_6 ~ Pct_publictrans_towork + pct_bachelor  + pct_white + pct_indian + pct_asian
        +pct_poverty + pct_onemployment + pct_pubtrans_native + pct_pubtrans_foreign + upper_rent + lower_rent,
            data=clean_trancts_2019_only,
              adapt = bw, #adaptive bandwidth determined by proportion of observations accounted for
              gweight=gwr.Gauss,
              se.fit=TRUE, #to return local standard errors
              hatmatrix = TRUE)
gwrmodel
summary(gwrmodel$SDF)
kappa( matrix, exact = FALSE)
```

